--- 
title: 'Problem Set #3'
output: 
  pdf_document: default
bibliography: "../bib/references.bib"
link_citations: true
csl: "../bib/apa.csl"
--- 

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, out.width = '75%', fig.align = 'center')
options(repos = list(CRAN="http://cran.rstudio.com/"))

# Load libraries
library(tidyverse)
library(scales)
library(knitr)
install.packages("kableExtra")
library(kableExtra)
install.packages('plyr', repos = "http://cran.us.r-project.org")

# Directory paths
data_dir <- file.path(".", "files")

```


# About the data
@housing describes this dataset as information regarding housing prices in a given district taken from the 1990 California Census [@housing]. 

```{r}
getwd()
list.files(data_dir)
housing_df <- read_rds(file.path(data_dir, "housing.RDS"))
 housing_df %>%
      head(10) %>% 
      kable() %>%
      kable_styling(latex_options = c('scale_down', 'HOLD_position'))
```


# Data analysis

#The homes inland cost the least money and are purchased more often by those with the lowest income.
```{r}
housing_df %>% ggplot(aes(x=median_income, y=median_house_value, color=ocean_proximity)) + geom_point() + ggtitle("Home Value by Income and Proximity to Ocean") + xlab("Income") + ylab("Home Value") + scale_color_brewer(palette = "Set3", name="Proximity to Ocean") + scale_y_continuous(labels = label_number(scale=1e-3, accuracy=1, suffix="K", prefix="$")) + scale_x_continuous(labels = label_number(scale=10, accuracy=1, suffix="K", prefix="$")) + theme_minimal()


```
#Most homes are not on an island.
```{r}
housing_df %>% ggplot(aes(x=median_income, y=median_house_value, )) + geom_point() + facet_wrap(~ ocean_proximity)  + ggtitle("Home Value by Income and Proximity to Ocean") + xlab("Income") + ylab("Home Value") + scale_color_brewer(palette = "Set3", name="Proximity to Ocean") + scale_y_continuous(labels = label_number(scale=1e-3, accuracy=1, suffix="K", prefix="$")) + theme_minimal()


```
#Homes by the ocean are between -125 and -117.5 degrees longitude.

```{r}
housing_df %>% ggplot(aes(x=longitude, y=latitude, color=ocean_proximity)) + geom_point() + ggtitle("latitude and longitude by ocean proximity") + xlab("Longitude") + ylab("Latitude") + scale_color_brewer(palette = "Set3", name="Proximity to Ocean") + coord_fixed() + theme_minimal()



```
#Homes by 40 degrees latitude are less expensive than 37.5 degrees latitude.

```{r}
housing_df %>% ggplot(aes(x=longitude, y=latitude, color=median_house_value)) + geom_point() + ggtitle("latitude and longitude by home value") + xlab("Longitude") + ylab("Latitude") + scale_color_gradient(low="white",high="purple", name="Home Value", labels=label_number(scale=1e-3, suffix="K", prefix="$")) + coord_fixed() + theme_minimal()


```
#Houses on the island are the most expensive.
```{r}
homevalue_count <- housing_df %>% group_by(ocean_proximity) %>% summarise(mean(median_house_value)) %>% rename("avg_home_value" = "mean(median_house_value)")

homevalue_count %>% ggplot(aes(x=ocean_proximity, y=avg_home_value)) + geom_col() + ggtitle("Home Value by Proximity to Ocean") + xlab("Proximity to Ocean") + ylab("Home Value") + scale_y_continuous(labels = label_number(scale=1e-3, accuracy=1, suffix="K", prefix="$")) + theme_minimal()


```
**Bonus Plot: This chart shows the split of the quantity of houses in terms of their distance to the ocean.** 
```{r}
pie(table(housing_df$ocean_proximity), main = "Count of People Living Near Water")
```


# References

